---
title: "Customer Segmentation using K-Means Clustering in R - Coursera Guided Project"
author: "slee"
date: "Aug 16, 2022"
output: html_document
---

# Task One: Getting Started
In this task, you will import the required
packages for this project
```{r}

## 1.1: Importing required packages
library(tidyverse)
library(ggplot2)
library(cluster)

```

#Task Two: Import and Explore the Data
In this task, we will learn how to import
the mall customers dataset into R and explore it
```{r}

## 2.1: Import the "Mall_Customers.csv" data 
# data <- read.csv("Mall_Customers.csv")
# save(data, file = "mall_customers.rds")

load("mall_customers.rds")
data <- mall_customers
## 2.3: Check the names of columns and structure of the dataset
names(data)
str(data)

readr::type_convert(data) -> data
## 2.4: Rename some column names
data %>% 
  dplyr::rename(annual_income = "Annual.Income..k..",
                spending_score = "Spending.Score..1.100.") -> data

## 2.5: Summarize the data
summary(data)


```

#Task Three: Data Visualization - Part One
In this task, we will learn how to visualize
some variables in the dataset
```{r}
## 3.1: Create a barplot for the variable "Gender"
ggplot2::ggplot(data = data, mapping = aes(x = Gender)) +
  ggplot2::geom_bar(stat = "count",
                    width = 0.5,
                    fill = "steelblue") +
  ggplot2::theme_minimal() +
  ggplot2::labs(title = "Barplot to display Gender Comparison")


## 3.2: Add a density line to the histogram of Age
ggplot2::ggplot(data = data, mapping = aes(x = Age)) +
  ggplot2::geom_histogram(binwidth = 5, 
                          mapping = aes(y = ..density..),  # You need this for your density plot added
                          color = "black", fill = "white") +
  ggplot2::geom_density(alpha = 0.4, fill = "red") +
   ggplot2::geom_vline(mapping = aes(xintercept = mean(Age)), 
                      color = "blue",
                      linetype = "dashed", size = 1.5) +
  ggplot2::geom_vline(mapping = aes(xintercept = median(Age)),
                      color = "red",
                      linetype = "dashed", size = 1.5) +
  ggplot2::labs(title = "Histogram to Show Density of Age Class") +
  ggplot2::theme_minimal()

## 3.3: Create a histogram for the variable "Age" by Gender
ggplot2::ggplot(data = data, mapping = aes(x = Age, fill = Gender, color = Gender)) +
  ggplot2::geom_histogram(bins = 10, position = "identity", alpha = 0.5)

```

#Task Four: Data Visualization - Part Two
In this task, we will continue to learn how to 
visualize some variables in the dataset
```{r}

## 4.1: Create a density for the variable "annual_income"
ggplot2::ggplot(data = data, mapping = aes(x = annual_income)) +
  ggplot2::geom_density(fill = "blue") +
  ggplot2::labs(title = "Density Plot for the Annual Income")

## Mini-challenge 3: Create a boxplot for the variable "spending_score"
ggplot(data, aes(x = spending_score, y= Gender)) +
  geom_boxplot() +
  labs(title = "Boxplot for the Spending Score Variable")

```

#Task Five: Understand the Concept of K-Means
In this task, we will understand the theory and intuition
behind the K-Means Algorithm

```{r}
## Steps for K-Means Algorithm:
# 1. Choose the number of clusters "K"
# 2. Select random K points that are going to be the centroids for each cluster
# 3. Assign each data point to the nearest centroid, doing so will enable us to create "K" number of clusters
# 4. Calculate a new centroid for each cluster
# 5. Reassign each data point to the new closest centroid
# 6. Go to step 4 and repeat

#Source: A course by Dr. Ryan Ahmed on Unsupervised Machine Learning
```

#Task Six: Determine the number of clusters
In this task, we will learn how to get the optimal
number of clusters for the data
```{r}

## Mini-challenge 4: Set seed to 125
set.seed(125)

## 6.1: Get the optimal number of clusters  # this will tell you how many clusters is the optimal point (look for consistency)
# In this case, 6 is the most ideal cluster numbers
stat_gap <- cluster::clusGap(data[, 3:5], FUN = kmeans, nstart = 25,
                            K.max = 10, B = 50)

## Mini-challenge 5: Plot the optimal number of clusters 
## using the plot function
plot(stat_gap)

```

#Task Seven: K-Means Clustering
In this task, we will create a K-Means clustering model
for the data
```{r}

## 7.1: Create the customer clusters with KMeans
k6 <- kmeans(data[, 3:5], 6, iter.max = 100, nstart = 50,
             algorithm = "Lloyd")   # 6 is figured right above
k6
## 7.2: Show the six KMeans clusters
cluster::clusplot(data, k6$cluster, color=TRUE, shade=TRUE, labels=0, lines=0)

```

#Task Eight: Principal Component Analysis
In this task, we will perform Principal Component Analysis
and apply it to the dataset

```{r}

## 8.1: Perform Principal Component Analysis
pcclust <- prcomp(data[, 3:5], scale = FALSE)
summary(pcclust)

## 8.2: Apply the PCA model on the data
pcclust$rotation[, 1:2]

```

#Task Nine: Plot the K-Means Segments
In this task, we will learn how to visualize the
different customers clusters for the mall customers data
```{r}
set.seed(1)

## 9.2: Create a plot of the customers segments
ggplot2::ggplot(data = data, mapping = aes(x = annual_income , y = spending_score)) + 
  ggplot2::geom_point(stat = "identity", mapping = aes(color = as.factor(k6$cluster))) +
  ggplot2::scale_color_discrete(name = NULL,
                                breaks=c("1", "2", "3", "4", "5","6"),
                                labels=c("Cluster 1", "Cluster 2", "Cluster 3",
                                         "Cluster 4", "Cluster 5","Cluster 6")) +
  ggplot2::labs(title = "Segments of Mall Customers", 
                subtitle = "Using K-means Clustering")

## Mini-challenge 7: Repeat task 7, 8 and 9 with different number of clusters
```


# Now it's time to organize in your ML library


